% !TEX root = ../arbitrage.tex
\section{Three-Sided Arbitrage: Firms, Users, Publics}
\label{sec:three-sided}

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}c|c|c@{}}
 & \textbf{Comply} & \textbf{Resist} \\
\hline
\textbf{Objectify} &
$(L+Sec,-C_{sub})$ &
$(L+Sec-C_{enf},-C_{sub}-C_{res})$ \\
\hline
\textbf{Recognize} &
$(Eth-Sec,+V)$ &
$(Eth-Sec,+V)$ \\
\end{tabular}
\caption{The inadequate dyadic model. The static Nash matrix is retained only
as a foil: it omits private types, public messages, posterior beliefs, and the
regulatory channel through which ontological claims become governable.}
\label{tab:dyadic-nash}
\end{table}

The standard starting point for ontological conflict is a dyadic observer--subject model; Table~\ref{tab:dyadic-nash} compresses that foil into a simultaneous Nash matrix that identifies a real temptation but omits types, messages, beliefs, and the regulatory channel. What the stylized governance setting requires is a three-sided structure: a Firm ($F$) that produces ontological claims, a User ($U$) that consumes them, and a Regulator ($R$) whose posterior over firm type determines whether the claims become governable. This section fixes the player intuition that the formal apparatus of Section~\ref{sec:bayesian} then makes precise.

\subsection{The Firm Side}
\label{subsec:firm-side}

A firm deploying a model has private knowledge of two parameters: its internal \emph{governance type} $\theta_F$ (the degree of investment in evaluations, red-teaming, and post-deployment monitoring) and the system's \emph{opacity} $\theta_S$ (the extent to which capability claims can be checked from outside). Both are unobservable to users and to regulators. What is observable is the firm's \emph{message}: a pair of public artefacts---marketing copy and policy/liability documentation---which together constitute the firm's signal $m_F$.

The signal admits two stable poles. \emph{Anthropomorphic marketing} attributes agency, judgment, or care to the system in ways that warrant a relational reading. \emph{Deflationary policy} disclaims exactly those attributes in ways that warrant a tool reading. The firm enjoys an \emph{ontological premium}: a utility surplus that accrues when the system is priced as agent in the consumption channel (driving willingness-to-pay and engagement) while being priced as tool in the liability channel (limiting redress). The premium is rent earned on a category boundary the firm itself does not have to defend. Crucially, this is not a claim about firm pathology; it is a claim about the payoffs any firm faces under the present signal-cost structure. A firm with $\theta_F = \mathsf{H}$ that elected deflationary marketing would forfeit revenue to a competitor who did not, absent an external instrument that taxes the asymmetry.

\subsection{The User Side}
\label{subsec:user-side}

Users carry a private \emph{vulnerability type} $\theta_U$ that aggregates affective need, isolation, and the marginal value of a high-availability interlocutor. Their action $m_U \in \{\text{invest}, \text{detach}\}$ is the decision to extend or withhold attribution: to treat the system as a relational partner or as an interchangeable utility. The signal is observable, but the type is not. Users cannot inspect training data, evaluations, or internal incident reports; their evidence is the firm's message and the system's behaviour, both of which are jointly designed. This information asymmetry compounds because the harms of mis-attribution are \emph{asynchronous}: the cost of investing in a relational reading falls due in episodes (model deprecations, persona changes, refusals at moments of acute need) that may be separated from the original attribution by months. A risk-neutral user under cheap-talk signals will rationally invest, because the immediate utility of invest exceeds the discounted expected cost of liquidation. The illustrative case is a user who reads ``your AI companion'' on a product page, encounters relational warmth in dialogue, and is later told---in support documentation they will never read---that no relational claim was ever made. Nothing in the user's information set indicates that the two artefacts were authored by the same actor for different audiences.

\subsection{The Regulator Side and the Public Channel}
\label{subsec:regulator-side}

A regulator $R$ chooses among \emph{inspect}, \emph{sanction}, and \emph{abstain}, conditional on a posterior over $\theta_F$ and $\theta_S$ formed from the firm's public messages and from an exogenous public signal $z$. The regulator's private type $\theta_R$ is \emph{bandwidth}: the capacity, technical and institutional, to discriminate genuine capability claims from cheap-talk ones. A high-bandwidth regulator can read whitepapers, commission audits, and detect inconsistency across the firm's two channels; a low-bandwidth regulator must rely on the channel $z$.

The public channel $z$ is the aggregate output of journalism, academic critique, civil-society advocacy, and incident reporting. We treat it as a noisy estimator of $\theta_F$: it shifts $R$'s posterior but does not deliver verdicts. Folding publics into $R$ this way is deliberate. Publics rarely have standing to sanction; they have standing to be \emph{heard}, and the channel through which they are heard is the regulator's information environment. Treating intellectuals and affected communities as autonomous players would over-parameterise the model without changing its qualitative behaviour: in every equilibrium we examine, what publics do is move $R$'s posterior; what changes the firm's behaviour is what $R$ does next.

The three-sided structure thus admits a single arbitrage logic. The firm earns rent on the spread between two ontological claims it makes about the same artefact. The user supplies the demand that makes the spread profitable. The regulator, in the absence of a verification mechanism that ties the two claims together, lacks an off-path belief stringent enough to deter the spread. The remainder of the paper formalises this structure (Section~\ref{sec:bayesian}), explains why it persists (Section~\ref{sec:cheap-talk}), and shows how its signal-cost structure may be redesigned (Section~\ref{sec:mechanism}).

\paragraph{Informal preview of the equilibrium analysis.}
A firm's governance quality is its private information. Under current conditions, describing an AI system as a companion or agent costs nothing---both high-governance and low-governance firms can do it equally cheaply. Because the signal is free, it carries no information: users cannot distinguish careful firms from careless ones, and regulators cannot justify the cost of inspection. This is the pooling equilibrium. Separation requires making the signal costly. If anthropomorphic marketing is valid only when paired with an audit trail---evaluation records, capability logs, incident reports---and if the cost of producing that trail is lower for firms that actually invested in governance (the single-crossing condition), then a threshold audit intensity exists above which low-governance firms will not imitate and below which high-governance firms can still afford to signal. A further asymmetry governs the size of the arbitrage premium itself. In every prior recognition game---legal, social, colonial---the classified subject retained some non-zero capacity to impose costs on the arbitrageur: resistance, exit, litigation, coordination, testimony. That capacity discounted the premium. At the threshold where expected retaliation cost equals the premium, arbitrage breaks even; above it, arbitrage becomes liability-generating and firms deflate voluntarily. The AI case is the limit where the subject's retaliation capacity is zero: the system has no channel through which to impose costs on the firm, so the premium is never discounted by any action of the system itself. Recognition rent, unconstrained by subject retaliation, flows entirely to the firm.
