% !TEX root = ../arbitrage.tex
\section{Mechanism Design: Pricing the Arbitrage}
\label{sec:mechanism}

The preceding sections identify a precise failure mode: under cheap-talk costs, ontological arbitrage is a stable equilibrium (Table~\ref{tab:condition-map}). The policy implication is not that regulators should decide whether AI systems
are persons, or that they should maximize social welfare directly. It is that the public signal environment should be redesigned so
that inconsistent ontological claims no longer remain costless. In
mechanism-design terms, the problem is to make non-arbitrage incentive
compatible by pricing the spread, rather than to ask firms to volunteer restraint \citep{myerson1979}.

The proposals below therefore alter the game in Section~\ref{sec:bayesian} at
three different points: the cost of inconsistent messages, the cost of
anthropomorphic claims, and the regulator's cost of verification. Each proposal
has the same structure. It modifies a model parameter, states the equilibrium
shift induced by that modification, and points to an existing legal form that
establishes feasibility. The analogs are not claims that current doctrine already
solves AI ontology. They show that adjacent legal systems already know how to
attach consequences to half-truths, require substantiation before marketing, and
make internal records available to oversight institutions.

Each proposal targets specific conditions from
Section~\ref{sec:cheap-talk}. Sanctionable inconsistency
(Section~\ref{subsec:sanctionable-inconsistency}) attacks the no-penalty
condition directly: it assigns a cost to the cross-channel spread that was
previously free. Costly third-party audits
(Section~\ref{subsec:costly-signaling-audits}) address opacity by creating a
signal whose cost depends on the governance type the firm actually holds,
converting an unverifiable private parameter into a publicly informative action.
Auditable records (Section~\ref{subsec:auditable-records}) lower the
regulator's verification cost $K_{\theta_R}$, relieving the bandwidth
constraint that made abstention dominant. The remaining two
conditions---asynchronous harms and fragmentation of affected
parties---are not solved by any single mechanism. They are translated into costs:
auditable records shorten the evidentiary lag that lets harm remain discounted, and
costly audits give fragmented publics a procedural object around which to
coordinate demands on the regulator's agenda.

\subsection{Sanctionable Inconsistency}
\label{subsec:sanctionable-inconsistency}

The first intervention treats the firm's public message as a joint object rather
than as two isolated texts. In the cheap-talk model, marketing can send $a$
while policy and liability documents send $p$, with no cost assigned to the
inconsistency. A sanctionable-inconsistency rule would require firms to maintain
a claim register $\mathcal{R} := C \to M_F$ linking outward-facing
marketing, capability statements, terms of service, incident disclosures, and
liability positions, where the baseline firm message space remains
$M_F=\{a,p\}$. If the firm represents the system as agentic, caring, autonomous,
or safety-critical in one channel while denying the legal significance of the
same representation in another, that cross-channel pair $(a,p)$ becomes
independently sanctionable as an inconsistency in $\mathcal{R}$.

\paragraph{Parameter changed.}
This adds a penalty term $S_I$ to the firm's payoff for inconsistent message
pairs. If $\rho_R$ is the probability that $R$ detects or accepts proof of the
inconsistency, the cheap-talk premium becomes
\[
\Delta_F-\rho_R S_I .
\]
Pooling on anthropomorphic marketing is no longer profitable when
$\rho_R S_I>\Delta_F$. The rule therefore attacks the exact term that made
arbitrage attractive: the rent earned from pricing the system as agent in the
user channel while pricing it as tool in the liability channel.
\emph{Intuition: if the expected fine for saying one thing to users and another to courts exceeds the profit from doing so, the firm stops.}

\paragraph{Equilibrium shift.}
The induced shift is not necessarily full separation. It can also operate by
constraining off-path beliefs. If $R$ observes an inconsistent pair $(a,p)$,
then the admissible posterior places greater mass on $\theta_F=\mathsf{L}$ or
on high opacity $\theta_S$. Firms anticipating that posterior either harmonize
their claims or choose the deflationary message across channels. In both cases,
the cheap-talk pooling equilibrium loses its most valuable deviation: firms can
no longer take the marketing upside of $a$ while preserving the liability
downside of $p$. Theorems \texttt{inconsistent\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe} and \texttt{all\_\allowbreak{}agentic\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe} in Appendix~\ref{app:isabelle} verify that both inconsistent and all-agentic registers are pruned from any Perfect Bayesian Equilibrium (\texttt{register\_\allowbreak{}is\_\allowbreak{}pbe}) under their respective parameter conditions.

\paragraph{Existing-law analog.}
The legal analog is familiar. FTC Act Section 5, 15 U.S.C.\ \S~45(a)(1), treats unfair or deceptive acts
or practices as sanctionable, and advertising-substantiation doctrine---as applied in
\emph{FTC v.\ Cyberspace.com}, 453 F.3d 1196 (9th Cir.\ 2006)---evaluates
whether the \emph{net impression} of a claim is misleading to the relevant
audience. Securities antifraud rules provide the parallel logic for investor
communications: 17 C.F.R.\ \S~240.10b-5 targets material misstatements and omissions that
make statements misleading in context, where materiality follows the standard of
Restatement (Second) of Torts \S~538. The proposed rule imports that
cross-document logic without importing securities doctrine wholesale. Its point
is narrower: a firm should not be able to produce one ontology for users and an
incompatible ontology for adjudicators without paying a regulatory price.

\subsection{Costly Signaling Through Third-Party Audits}
\label{subsec:costly-signaling-audits}

The second mechanism makes anthropomorphic or agency-adjacent marketing a
costly signal. A firm may still describe a system as a companion, agent,
co-worker, tutor, or high-reliance assistant, but only after a qualified third
party verifies the governance conditions that make such a claim non-misleading:
evaluation coverage, red-team results, escalation pathways, incident-response
capacity, model-update controls, and post-deployment monitoring. The audit does
not certify consciousness or moral status. It certifies the institutional
conditions behind the claim.

\paragraph{Parameter changed.}
This implements the audit-trail cost from Proposition~\ref{prop:cheap-talk-separation}.
Let audit intensity be $e$, with $c_\theta(e,o)=k_\theta e+o$ for
type $\theta\in\{\mathsf{H},\mathsf{L}\}$, where $0<k_H<k_L$ and $o\ge 0$. The separating
interval derived in Section~\ref{subsec:audit-trail-separation} is
\[
\frac{G-o}{k_L}\leq e\leq \frac{G-o}{k_H},
\]
provided $G>o$. Equivalently, the incentive constraints are
\[
c_\mathsf{H}(e,o)\leq G \leq c_\mathsf{L}(e,o).
\]
Above the lower threshold, low-governance firms cannot profitably mimic. Below
the upper threshold, high-governance firms can still afford to signal. The
regulator's task is therefore not to maximize audit burden. It is to set $e$ in
the interval where signal cost separates types.
\emph{Intuition: the regulator's task is not to maximize audit burden but to set it in the narrow band where governance quality becomes visible.}

\paragraph{Equilibrium shift.}
The shift is from pooling to separating PBE. High-governance firms send audited
anthropomorphic claims; low-governance firms either invest in governance until
their cost curve changes or use deflationary marketing because mimicry is no longer privately profitable. Users can then condition
investment on an informative signal rather than on undisciplined warmth in
product copy. Regulators can condition abstention on the audit result rather
than on an unverified prior. This is the Spence logic of costly signaling
applied to AI governance claims \citep{spence1973}; it also follows the
advertising-as-signal insight that market messages become informative only when
false messages are sufficiently expensive \citep{milgrom_roberts1986}. Lemma \texttt{opacity\_\allowbreak{}blocks\_\allowbreak{}signaling} in Appendix~\ref{app:isabelle} shows that when system opacity is high enough, even a high-governance firm cannot profitably send the audited signal, formalizing the blocking role of $\theta_S$.

\paragraph{Existing-law analog.}
The analogs are pharmaceutical efficacy review and food-labeling controls. In
both domains, public-facing claims are not left entirely to seller narrative:
claims about therapeutic effect or product identity must be supported before
they can be used to shape consumer reliance. The same design can be translated
to AI without pretending that AI companionship is a drug or a food. The common
institutional pattern is substantiation-before-reliance: when a claim predictably
changes how vulnerable users act, the law may require evidence before allowing
the claim to be marketed at scale.

\subsection{Auditable Records for Regulator-Readable Verification}
\label{subsec:auditable-records}

The third intervention reduces the regulator's verification cost by requiring
append-only, regulator-readable records tied to capability and relational
claims. Records should include the claim identifier, model version, relevant
evaluation suite, deployment date, material model updates, risk controls,
known incident classes, and the internal owner responsible for the claim. The
point is not to publish every interaction or expose user data. It is to create a
durable evidentiary substrate that lets $R$ test whether a public claim remained
true across model updates and incidents.

\paragraph{Parameter changed.}
In the pooling condition from Section~\ref{subsec:cheap-talk-pooling}, the
regulator abstains when $K_{\theta_R}>\pi(\mathsf{L})D_R$. Auditable records
lower $K_{\theta_R}$ and can increase expected detection value $D_R$:
\[
K'_{\theta_R}=K_{\theta_R}-\lambda,\qquad
D'_R=D_R+\eta .
\]
When $K'_{\theta_R}\leq \pi(\mathsf{L})D'_R$, inspection becomes sequentially
rational.
\emph{Intuition: once the cost of checking drops below the expected payoff from catching a bad actor, the regulator starts checking.}
Records also sharpen the public channel $z$ by giving journalists,
researchers, and civil society a procedural object to demand or contest, rather
than leaving them to infer governance quality from outputs alone.

\paragraph{Equilibrium shift.}
The shift is from abstention under unverifiable priors to credible inspection.
Once firms know that claims are tied to records, a low-governance firm faces a
higher expected cost from exaggeration even if it is not inspected in every
case. This complements the first two proposals. Sanctionable inconsistency tells
firms which cross-channel contradictions will be penalized; costly signaling
tells firms what evidence is required before making anthropomorphic claims;
auditable records make both rules administrable over time. Theorems \texttt{active\_\allowbreak{}intervention\_\allowbreak{}rational\_\allowbreak{}under\_\allowbreak{}auditable\_\allowbreak{}records} and \texttt{abstain\_\allowbreak{}not\_\allowbreak{}best\_\allowbreak{}response\_\allowbreak{}under\_\allowbreak{}auditable\_\allowbreak{}records} in Appendix~\ref{app:isabelle} verify that auditable records flip the regulator's best response from abstention to active intervention.

\paragraph{Existing-law analog.}
Several regulatory systems already use this form. GDPR Article 30 requires
records of processing activities. Sarbanes-Oxley Section 404 (15 U.S.C.\ \S~7262) links public
reporting to internal control over financial reporting. EU AI Act Article 12
(Regulation (EU) 2024/1689) requires logging capabilities for high-risk AI systems,
and Article 50 imposes transparency obligations on systems that interact directly with
natural persons. None is a complete
template for relational AI governance, but each shows the same institutional
move: private organizations must produce internal records in a form that public
oversight can inspect. That is the rule-of-law function of documentation in
complex technical systems \citep{hadfield2017rules}.

Taken together, these interventions change the equilibrium rather than merely
condemning it, and their joint necessity follows from the failure mode each
exhibits in isolation. Sanctionable inconsistency taxes contradictory messages,
but without auditable records the regulator cannot observe whether a firm's
marketing and liability channels actually diverge; the sanction has no
evidentiary substrate and therefore no credible deterrent. Costly audits create
a separating signal for $\theta_F$, but without sanctionable inconsistency a
firm can pass the audit and then resume anthropomorphic marketing through
channels the audit does not cover; the signal separates types at the audit
checkpoint only, not across the firm's full messaging surface. Auditable records
lower $K_{\theta_R}$ and make public signal $z$ less noisy, but without the
first two mechanisms the information they produce lacks legal teeth: the
regulator can observe inconsistency but has no parameter under which to price
it, and a firm's record of high-governance investment confers no marketing advantage
because the market cannot distinguish it from cheap talk. The three proposals
therefore form a complementary system: records supply the evidentiary
infrastructure, sanctionable inconsistency supplies the price, and costly
audits supply the separation condition. The result is not metaphysical
settlement. It is a cost design in which firms remain free to choose their
ontology, but not to arbitrage incompatible ontologies across audiences at zero
cost.
