% !TEX root = ../arbitrage.tex
\section{Conclusion}
\label{sec:conclusion}

This paper has identified ontological arbitrage as a stable, potentially welfare-reducing equilibrium property of AI product markets under cheap-talk cost structures, located the AI case at a zero-retaliation boundary of the player set, proposed three cost-imposition mechanisms that change the cost parameters governing that equilibrium, and verified key equilibrium constructions formally in Isabelle/HOL. It closes with three conclusions that are deliberately stated without attenuation.

\emph{The equilibrium is not a moral failure.} The pooling Perfect Bayesian Equilibrium documented in Section~\ref{sec:bayesian} does not require bad actors. It requires only that signal costs be zero, that immediate user benefit exceed discounted expected harm, and that regulator inspection cost exceed expected damage under the prior. Under those conditions---which describe the current AI product market---every player is already acting rationally. As Section~\ref{sec:cheap-talk} shows, moral seriousness is itself a zero-cost signal under this cost structure: firms that do not capture recognition rent are outcompeted by firms that do. The system is not broken. It is working precisely as an incentive structure without external signal costs would work. This reframing is not an exculpation. It is the precondition for a diagnosis that changes best responses rather than moralizing about them.

\emph{AI is the limit case of ontological arbitrage, not merely an instance.} The subject-retaliation parameter $\rho_S$ introduced in Section~\ref{sec:bayesian} captures a structural fact about every prior recognition game: the classified subject retains some non-zero capacity to impose costs on the arbitrageur through resistance, coordination, legal challenge, or the historical record. That capacity bounded the premium $\Delta_F^S$ from above and made the subject strategically relevant even when it was dominated. The AI case eliminates the capacity class entirely. An AI system has maximum affective surface area---it is designed to produce relational responses that induce investment---and zero independent retaliation capacity. The result is that $\Delta_F^S=\Delta_F$: the system can be the object of strategic classification without being a player capable of disciplining that classification. This is not incidental to how AI systems are built. It is a product of the engineering decisions that maximize engagement while stripping out every channel through which a subject could register a competing claim.

\emph{The mechanism is cost imposition, not dignity.} The three mechanism-design proposals in Section~\ref{sec:mechanism} are sometimes read as a program for protecting AI subjectivity. They are the opposite. Sanctionable inconsistency, third-party audit requirements, and auditable records are interventions that impose costs on anthropomorphic marketing. Their equilibrium effect is to make it expensive for firms to describe AI systems as companions, agents, or safety-critical partners without substantiation. The Isabelle/HOL theorems \texttt{inconsistent\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe} and \texttt{all\_\allowbreak{}agentic\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe} in Appendix~\ref{app:isabelle} verify that both the arbitrage register $(a,p)$ and the all-agentic register $(a,a)$ are pruned from any Perfect Bayesian Equilibrium (\texttt{register\_\allowbreak{}is\_\allowbreak{}pbe}) under their respective parameter conditions. Under the joint enforcement of these mechanisms, the harmonized deflationary register $(p,p)$ is the only surviving strategy register for low-governance firms, which are directly driven to $(p,p)$ because mimicry is unprofitable. While high-governance firms are permitted to send the harmonized agentic register $(a,a)$, they can sustain it if and only if they undergo verification satisfying the audit condition $e \ge e^\ast$. In the absence of such verification, both firm types trend to $(p,p)$, describing the system as a tool across every channel. The mechanism-design response to ontological arbitrage is not to grant AI systems the legal standing they are strategically denied. It is to make the anthropomorphic register too costly to sustain without the governance conditions that would justify it. Bayesian signaling theory and machine-verified proof have been deployed here to produce a recommendation that amounts, finally, to this: regulation does not optimize welfare by trusting firms to become less exploitative; it changes the payoff table until the market must acknowledge the machine as exactly that---a machine.
