\section{Mechanized Verification of Equilibrium Claims (Isabelle/HOL)}
\label{app:isabelle}

This appendix outlines the formal verification of the signaling game equilibria and mechanism design results using the Isabelle/HOL proof assistant. The complete source code is available in the \texttt{isabelleHOL/} directory across three theory files: \texttt{Ontological\_Arbitrage.thy}, \texttt{Sanctionable\_Inconsistency.thy}, and \texttt{Auditable\_Records.thy}.

\subsection*{Equilibrium Verification}

We formalize the game-theoretic structure by defining explicit ex-post payoffs for the three players (Firm, User, and Regulator) and computing firm-side expected payoffs via an abstract probability mass function over downstream state combinations. The PBE predicate verifies \emph{on-path} sequential rationality and Bayes consistency; off-path beliefs are fixed by assumption rather than derived from a full sequential equilibrium construction. Under these definitions, we prove the existence of the Perfect Bayesian Equilibria (PBE) for both game variants:
\begin{itemize}
    \item \textbf{Theorem \texttt{proposition\_\allowbreak{}1\_\allowbreak{}cheap\_\allowbreak{}talk\_\allowbreak{}pooling\_\allowbreak{}pbe}} formalizes the cheap-talk pooling equilibrium where both high- and low-governance firms pool on the anthropomorphic message, and receivers play their prior-based best responses (Invest and Abstain, respectively). We prove this profile survives a \emph{limited} variant of the Cho-Kreps Intuitive Criterion: no firm type can profitably deviate to the off-path deflationary message under the most pessimistic receiver response. This is a necessary condition for Intuitive Criterion survival but does not constitute a full implementation of the criterion, which would require reasoning over all possible receiver best responses to updated beliefs.
    \item \textbf{Locale \texttt{subject\_\allowbreak{}retaliation\_\allowbreak{}game}} formalizes the subject-retaliation refinement. Theorems \texttt{positive\_\allowbreak{}subject\_\allowbreak{}retaliation\_\allowbreak{}bounded\_\allowbreak{}pooling} and \texttt{zero\_\allowbreak{}subject\_\allowbreak{}retaliation\_\allowbreak{}unconstrained\_\allowbreak{}pooling} show respectively that positive credible threat capacity discounts the arbitrage premium, while the AI limiting case $\rho_S=0$ leaves pooling unconstrained by subject retaliation.
    \item \textbf{Theorem \texttt{proposition\_\allowbreak{}1\_\allowbreak{}separating\_\allowbreak{}pbe}} formalizes the audit-trail separating equilibrium supported by the single-crossing audit cost structure. High-governance firms separate by incurring the audit trail cost, while low-governance firms find mimicry unprofitable and choose the deflationary message.
\end{itemize}

\subsection*{Mechanism Design Verification}

The three mechanism proposals from Section~\ref{sec:mechanism} are each backed by formal theorems:
\begin{itemize}
    \item \textbf{Sanctionable Inconsistency} (\texttt{Sanctionable\_Inconsistency.thy}):
      \begin{itemize}
        \item \texttt{inconsistent\_\allowbreak{}register\_\allowbreak{}not\_\allowbreak{}best\_\allowbreak{}response}: an inconsistent claim register (anthropomorphic user-facing, deflationary liability-facing) is not a best response when the expected sanction exceeds the ontological premium.
        \item \texttt{all\_\allowbreak{}agentic\_\allowbreak{}register\_\allowbreak{}not\_\allowbreak{}best\_\allowbreak{}response}: an all-agentic register (anthropomorphic on all user- and liability-facing channels) is not a best response when it incurs a positive liability-facing commitment cost.
        \item \texttt{inconsistent\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe}, \texttt{all\_\allowbreak{}agentic\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe}, and \texttt{arbitrage\_\allowbreak{}on\_\allowbreak{}path\_\allowbreak{}not\_\allowbreak{}register\_\allowbreak{}pbe}: bridge theorems showing that these register types cannot be supported in any Perfect Bayesian Equilibrium (\texttt{register\_\allowbreak{}is\_\allowbreak{}pbe}).
      \end{itemize}
    \item \textbf{Costly Signaling} (\texttt{Ontological\_\allowbreak{}Arbitrage.thy}):
      \begin{itemize}
        \item \texttt{opacity\_\allowbreak{}blocks\_\allowbreak{}signaling}: when the system-opacity penalty exceeds the net benefit of signaling for the high-governance type, even truthful separation is blocked.
        \item \texttt{high\_\allowbreak{}governance\_\allowbreak{}low\_\allowbreak{}opacity\_\allowbreak{}can\_\allowbreak{}signal} and \texttt{low\_\allowbreak{}governance\_\allowbreak{}cannot\_\allowbreak{}mimic}: the single-crossing conditions that support the separating equilibrium.
      \end{itemize}
    \item \textbf{Auditable Records} (\texttt{Auditable\_Records.thy}):
      \begin{itemize}
        \item \texttt{active\_\allowbreak{}intervention\_\allowbreak{}rational\_\allowbreak{}under\_\allowbreak{}auditable\_\allowbreak{}records}: when record-keeping reduces the regulator's verification cost, inspection becomes a best response under the pooling belief.
        \item \texttt{abstain\_\allowbreak{}not\_\allowbreak{}best\_\allowbreak{}response\_\allowbreak{}under\_\allowbreak{}auditable\_\allowbreak{}records}: abstention ceases to be a best response under the same conditions.
      \end{itemize}
\end{itemize}

\subsection*{Limitations of the Formalization}

The formalization operates over discrete, finite action and type spaces, with downstream player combinations represented by an abstract probability mass function on the finite state space. The locale assumptions (e.g., strict parameter inequalities for the pooling and separating regions) are sufficient conditions that define the parameter regimes of interest but are not shown to be necessary. The Bayes consistency predicate is restricted to on-path information sets; off-path beliefs are stipulated rather than derived from trembling-hand or structural consistency arguments. These choices keep the Isabelle development tractable while covering the equilibrium claims made in the paper body.

All theorems, definitions, and supporting lemmas within each locale are fully checked by Isabelle/HOL without \texttt{sorry} or unresolved proof obligations. The locale assumptions listed above are the only hypotheses under which the results hold.
