\documentclass[letterpaper]{article}

\usepackage[submission]{aaai2026}
\setcounter{secnumdepth}{2}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{natbib}

\usetikzlibrary{positioning,calc}

\urlstyle{rm}
\def\UrlFont{\rm}
\frenchspacing
\ifdefined\pdfpagewidth
  \setlength{\pdfpagewidth}{8.5in}
  \setlength{\pdfpageheight}{11in}
\fi

\newtheorem{proposition}{Proposition}

\title{Recognition Rent at the Zero-Retaliation Limit}

\author{
    Anonymous Submission
}
\affiliations{
    Paper under review
}

\begin{document}

\maketitle

\begin{abstract}
Game-theoretic modeling begins by fixing the player set. This paper studies a market in which that boundary is itself a source of surplus: firms sell recognition-like claims about AI systems that cannot retaliate against the classifications imposed on them. Ontological arbitrage arises when a firm describes its AI system as a relational agent to users---driving willingness-to-pay and engagement---while disclaiming agency in terms of service to limit liability. We call the resulting surplus \emph{recognition rent}: the premium a firm captures when the represented object has no retaliatory channel through which recognition can become costly. The model is not a welfare-optimization account: rent extraction may reduce aggregate welfare while remaining individually rational, because the firm internalizes the premium and externalizes delayed harm. We model this behavior as a sequential signaling game under incomplete information within a three-sided Firm-User-Regulator structure. A subject-retaliation parameter $\rho_S$ captures the key structural variable: setting $R(\rho_S) = \Delta_F - \rho_S C_S$, every prior recognition game operated under $\rho_S > 0$, where the classified subject's capacity to resist bounded the premium from above. The AI case is the limit $\rho_S = 0$: the system is the object of strategic classification but not a strategic player capable of imposing off-equilibrium costs. Under cheap-talk signaling costs, a pooling Perfect Bayesian Equilibrium exists in which both governance types employ anthropomorphic marketing, leaving posteriors equal to priors. In that region, unilateral non-exploitation is not a best response. Under a single-crossing audit-trail cost, a separating equilibrium emerges. We propose three cost-imposition mechanisms---sanctionable inconsistency, costly third-party audits, and auditable records---that jointly convert anthropomorphic marketing from cheap talk into a costly signal. Key equilibria are verified in Isabelle/HOL with zero \texttt{sorry} obligations.
\end{abstract}

\input{../sections/01_introduction}
\input{../sections/02_related_work}
\input{../sections/04_three_sided}
\input{../sections/03_bayesian_signaling}
\input{../sections/07_cheap_talk_persists}
\input{../sections/08_mechanism_design}
\input{../sections/09_conclusion}

\appendix
\input{../sections/A_isabelle_appendix}

\bibliography{arbitrage}

\end{document}
